State of the Practice for Lattice Boltzmann Method Software

### 2022-04-20 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Edits up to and including Section 11 (Threats to Validity)

	\item[Verifiability] An artifact (like code, or a theory manual) is
	verifiable if, and only if, every statement therein is verifiable. A
	statement is verifiable if, and only if, there exists some finite
	cost-effective process with which a person or machine can check that each
	statement is correct (adapted from recommended practice for requirements
	specifications \citep{IEEE1998}). Similarly to correctness, verifiability is
	correlated with the availability of a specification and with reference to
	domain knowledge. A good measure of verifiability is further correlated with
	the availability of well written tutorials that include expected output,
	with software unit testing, and with evidence of continuous integration. In
	our process we measure correctness and verifiability together.

@@

The use of continuous integration tools and techniques alludes to a more refined
development process where faults are isolated and better recognized. Only three
of the packages (ESPResSo, Ludwig, Musubi) mentioned applying the practice of
continuous integration in their development process.

@@

Figure~\ref{Fig_CorrectnessVerifiability} shows the ranking for surface
correctness and verifiability. Software packages with a higher score tend to
have theory documentation. They also explicitly use at least one document
generation tool that helps with verification, thus building confidence in
correctness. The top ranked software packages all include an easy to follow
getting started tutorial, and most of these include expected output. The two top
ranked packages, Ludwig and Musubi, indicate the use of unit testing in their
documentation. They also incorporated continuous integration in the development
process. Furthermore, eight of the top 10 ranked packages are noted as being
alive.

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{ p{2.5cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1.2cm}p{1cm}p{0.8cm} }
\toprule
~ \ & \cite{USGS2019} & \cite{TobiasEtAl2018} & \cite{BrettEtAl2021} &
\cite{WilsonEtAl2016} & \cite{SmithAndRoscoe2018} & \cite{HerouxEtAl2008} &
\cite{ThielEtAl2020} & \cite{vanGompelEtAl2016} & \cite{OrvizEtAl2017} & LBM\\
\midrule
LICENSE & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & C\\
README &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & C\\
CONTRIBUTING &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & U\\
CITATION &  &  &  & \checkmark & & & & \checkmark & \checkmark & U\\
CHANGELOG &  & \checkmark &  & \checkmark & \checkmark & & \checkmark &  &  & U\\
INSTALL &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & C\\
\midrule
Uninstall &  &  &  &  & & & & \checkmark & & R\\
Dependency List &  &  & \checkmark & & \checkmark & & & \checkmark &  & C\\
Authors &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & C\\
Code of Conduct &  &  &  &  & & & \checkmark & & & \\
Acknowledgements &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & R\\
Code Style Guide &  & \checkmark &  &  & & & \checkmark & \checkmark & \checkmark & R\\
Release Info. &  & \checkmark &  &  & & \checkmark & \checkmark & & & U\\
Prod.\ Roadmap &  &  &  &  & & \checkmark & \checkmark & \checkmark & & R\\
\midrule
Getting started &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & C\\
User manual &  &  & \checkmark &  & & & \checkmark & & & U\\
Tutorials &  &  &  &  & & & \checkmark & & & C\\
FAQ &  &  &  &  & & & \checkmark & \checkmark & \checkmark & R\\
\midrule
Issue Track &  & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & & \checkmark & C\\
Version Control &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & C\\ 
Build Scripts &  & \checkmark &  & \checkmark & \checkmark & \checkmark &
\checkmark & & \checkmark & C\\
\midrule
Requirements &  & \checkmark &  &  & & \checkmark &  &  & \checkmark &  \\
Design Doc.\ &  & \checkmark  & \checkmark &  & \checkmark & & \checkmark &
\checkmark& \checkmark & U\\
API Doc. &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
Test Plan &  & \checkmark &  &  & & \checkmark & & & &  U\\
Test Cases & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & U\\
\bottomrule
\end{tabular}
\caption{Comparison of Recommended Artifacts in Software Development Guidelines
to Artifacts in LBM Projects (C for Common, U for Uncommon and R for Rare)}
\label{Tbl_Guidelines}
\end{center}
\end{table}

@@

\item \textbf{API documentation} is \wss{fill in definition, examples and
benefits.  talk about doxygen to support.}

@@

Development tools support the development of end products, but do not become
part of them, unlike dependencies that remain in the application once it is
released \citep[p.\ 506]{GhezziEtAl2003}. Although not shown in
Table~\ref{tbl_tools}, debuggers were also likely used.  Only three (ESPResSo,
Ludwig and Musubi) packages mentioned continuous integration tools, like Travis
CI. Code editors and compilers were explicitly noted to have been used by
several packages, and were likely used by all of them. One of the packages
(Ludwig) explicitly noted the use of proprietary unit testing code written in C.
Likewise, the use of proprietary code for verifying the correctness of output
was noted by one of the developers (pyLBM). Similar tools were likely used when
developing the other software packages.

@@

@@

As mentioned in Section~\ref{SecSurfCorrectAndVerifiab}, continuous integration
is rarely used in LBM software (3 of 24 packages or 12.5\%). This contrasts with
the frequency with which continuous integration is recommended in research
software development guidelines \citep{BrettEtAl2021, Brown2015, ThielEtAl2020,
Zadka2018, vanGompelEtAl2016}.  \wss{why continuous integration.}  We could not
find published data on the frequency with which continuous integration is used
in general research software.  \wss{add reference to Carver paper.}  Our
impression is that like LBM software, other research software packages lag
behind the recommended best practice of employing continuous integration.

Interviews with the developers revealed a potentially more frequent use of both
unit testing and continuous integration, compared to what was observed from
studying the repository artifacts.

@@

Interviews with developers confirmed a similar project management process. In
teams of only a couple of developers, additions of new features or major changes
are discussed with the entire team. Projects with more than a couple developers
have lead developer roles. These lead developers review potential additions to
the software. One of the developers (ESPResSo) that was interviewed noted that
an ad hoc peer review process is used to assess major changes and additions.
Using peer review (also called technical review) matches with recommended
practice for research software \citep{HerouxEtAl2008, Givler2020, OrvizEtAl2017,
USGS2019}.

@@

@@

\wss{how to doc motivations and design principles. examples. advantages.}

@@

@@

	\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
	\textbf{Technology Hurdles} Technology pain points for LBM developers
	include setting up parallelization and continuous integration. The pain
	point survey of \citep{WieseEtAl2019} also highlighted technical-related
	problems like dependency management, cross-platform compatibility,
	continuous integration, hardware issues and operating system issues.

@@

Some LBM developers improve reproducibility via automating the verification
tests and making them part of a continuous integration process (mentioned in
Section~\ref{Sec_CompareTools}).  Providing unit tests improves reproducibility
because future developers will require a means to verify that they can obtain
the same computational results as today.  Unit tests are a good way to embed
self-diagnostics of reproducibility in the code \citep{BenureauAndRougier2017}.

@@

In this section we provide recommendations to address the pain points from
Section~\ref{painpoints} to answer~\rqref{RQ_Recommend}.  Our recommendations
are not lists of criticisms for what should have been done in the past, or what
should be done now; they are just suggestions for consideration in the future.
We will not be repeating ideas here that have been discussed in previous
sections, such as continuous integration, documentation of APIs, etc.  Our aim
is to mention ideas that are at least somewhat beyond conventional best
practices. The ideas listed here have the potential to become best practices in
the medium to long-term. The ideas are presented roughly in the order of
increasing novelty.

@@

With the exception of \citet{ThielEtAl2020}, the research software guidelines
that we consulted do not mention linters.  However, we feel linters have the
potential to improve code quality at a relatively low cost.  A linter is a tool
that statically analyzes code to find programming errors, stylistic
inconsistencies and suspicious constructs \citep{Wikipedia2022_Lint}. Linters
can be used to spot check code files, or even better as part of a continuous
integration system.

@@

The threats to validity can be categorized into methodology and data collection
issues.  With respect to our methodology, the measurement template may not be
detailed enough to accurately capture some qualities. For example, there are
only two measures of surface robustness. The measurement of robustness could be
expanded, as it currently only measures unexpected input. Other intentional
faults could be introduced to assess robustness, but this would require a larger
investment of measurement time. Similarly, reusability
(Section~\ref{reusabilityresults}) is assessed via the number of code files and
LOC per file. While this measure is indicative of modularity, it is possible
that some packages have many files, with few LOC, but the files do not contain
source code that is easily reusable. The files may be poorly formatted, or the
source code may be vague and have ambiguous identifiers. Maintainability may
also be incorrectly estimated, since it is assumed to be improved by a higher
ratio of comments to source code (Section~\ref{Sec_Maintainability}), but this
is not necessarily true. Furthermore, the measurement of understandability
(Section~\ref{Sec_SurfUnderstandability}) relies on 10 random source code files.
By random chance, the 10 files that were chosen may not be a good representation
of the understandability of that package.

In our methodology, one individual measures each package and the beginning and
ending of the measurement phase is separated by several months.  There is a
threat that the measurement may be affected by unconscious biases, or the
reviewers judgement may drift over time.  In previous work
\citep{SmithEtAl2016}, we demonstrated that the measurement process is
reasonably reproducible by having five products graded by a second reviewer. The
ranking via this independent review was almost identical to the original
ranking. As long as each grader uses consistent definitions, the relative
comparisons in the AHP results will be consistent between graders.

Regarding data collection, a risk to the validity of this assessment is missing
or incorrect data. Some data may not have been measured due to technology issues
like broken links. This issue arose with the measurement of Palabos, which had a
broken link to its user manual, as noted in Section~\ref{Sec_CompareArtifacts}.
Some pertinent data may not have been specified in public artifacts, or may be
obscure within an artifact or web-page. For instance, the use of unit testing
and continuous integration was mentioned in the artifacts of only three
(ESPResSo, Ludwig, Musubi) packages (Section~\ref{SecSurfCorrectAndVerifiab}).
However, interviews suggested a more frequent use of both unit testing and
continuous integration. For example, OpenLB, pyLBM, and TCLB use such methods
during development despite this not being explicit from an analysis of the
material available online. Given that we could not interview a representative
for every project, to keep the measures consistent we did not modify our initial
repository inspection based CI count.

Furthermore, design documentation was measured to be a ``less common'' artifact
in this assessment, but it is probable that such documentation is part of all
LBM packages. After all, developing research software is not a trivial
endeavour. It is likely that many packages have such documentation but did not
make it public, and due to this the measured data may not be a true reflection
of software package quality.

Another data collection concern is the gap between the measurement of Musubi and
the measurement of all other packages (Section~\ref{identifysoftware}).
Unfortunately we were not aware of the existence of Musubi until after the
primary data collection period was completed. This gap has a secondary effect
because it caused a mismatch between the time of manual and automated data
collection (Section~\ref{repmetrics}).  With time and resource constraints we
could not redo the manual measurement, but we did update the automated measures
(like the number of files, LOC, etc.) when Musubi was added.

Our comparison between our ranking and the community's ranking
(Section~\ref{repmetrics}) may not be valid.  Using stars to estimate the
community's ranking is not necessarily accurate.  Moreover, our overall AHP
ranking makes the unrealistic assumption of equal weighting between qualities
(Section~\ref{Sec_OverallQuality}).

A final threat to validity is conceptual.  Our work assumes that following best
practices leads to improved software quality.  As the advice in software
engineering guidelines show, we are not alone in this assumption.  Ideally
software qualities should be measured directly and then the impact of best
practices on those qualities should be assessed.  However, researchers in this
domain always face the challenge that many qualities, like maintainability,
reusability and verifiability, are difficult to measure directly.  Therefore, we
follow the usual approach of assuming that following procedures and adhering to
standards yields a higher-quality product \citep[p.\ 112]{VanVliet2000}.

@@

TO do
- check references

>>>>

	\item[Verifiability] An artifact (like code, or a theory manual) is
	verifiable if, and only if, every statement therein is verifiable. A
	statement is verifiable if, and only if, there exists some finite
	cost-effective process with which a person or machine can check that each
	statement is correct (adapted from recommended practice for requirements
	specifications \citep{IEEE1998}). Similarly to correctness, verifiability is
	correlated with the availability of a specification and with reference to
	domain knowledge. A good measure of verifiability is further correlated with
	the availability of well written tutorials that include expected output,
	with software unit testing, and with evidence of Continuous Integration
	(CI). In our process we measure correctness and verifiability together.

@@

The use of Continuous Integration (CI) tools and techniques alludes to a more
refined development process where faults are isolated and better recognized.
Only three of the packages (ESPResSo, Ludwig, Musubi) mentioned applying the
practice of CI in their development process.

@@

Figure~\ref{Fig_CorrectnessVerifiability} shows the ranking for surface
correctness and verifiability. Software packages with a higher score tend to
have theory documentation. They also explicitly use at least one document
generation tool that helps with verification, thus building confidence in
correctness. The top ranked software packages all include an easy to follow
getting started tutorial, and most of these include expected output. The two top
ranked packages, Ludwig and Musubi, indicate the use of unit testing in their
documentation. They also incorporated CI in the development
process. Furthermore, eight of the top 10 ranked packages are noted as being
alive.

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{ p{2.5cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1.2cm}p{1cm}p{0.8cm} }
\toprule
~ \ & \citet{USGS2019} & \citet{TobiasEtAl2018} & \citet{BrettEtAl2021} &
\citet{WilsonEtAl2016} & \citet{SmithAndRoscoe2018} & \citet{HerouxEtAl2008} &
\citet{ThielEtAl2020} & \citet{vanGompelEtAl2016} & \citet{OrvizEtAl2017} &
LBM\\
\midrule
LICENSE & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & C\\
README &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & C\\
CONTRIBUTING &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & U\\
CITATION &  &  &  & \checkmark & & & & \checkmark & \checkmark & U\\
CHANGELOG &  & \checkmark &  & \checkmark & \checkmark & & \checkmark &  &  & U\\
INSTALL &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & C\\
\midrule
Uninstall &  &  &  &  & & & & \checkmark & & R\\
Dependency List &  &  & \checkmark & & \checkmark & & & \checkmark &  & C\\
Authors &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & C\\
Code of Conduct &  &  &  &  & & & \checkmark & & & \\
Acknowledgements &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & R\\
Code Style Guide &  & \checkmark &  &  & & & \checkmark & \checkmark & \checkmark & R\\
Release Info. &  & \checkmark &  &  & & \checkmark & \checkmark & & & U\\
Prod.\ Roadmap &  &  &  &  & & \checkmark & \checkmark & \checkmark & & R\\
\midrule
Getting started &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & C\\
User manual &  &  & \checkmark &  & & & \checkmark & & & U\\
Tutorials &  &  &  &  & & & \checkmark & & & C\\
FAQ &  &  &  &  & & & \checkmark & \checkmark & \checkmark & R\\
\midrule
Issue Track &  & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & & \checkmark & C\\
Version Control &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & C\\ 
Build Scripts &  & \checkmark &  & \checkmark & \checkmark & \checkmark &
\checkmark & & \checkmark & C\\
\midrule
Requirements &  & \checkmark &  &  & & \checkmark &  &  & \checkmark &  \\
Design Doc.\ &  & \checkmark  & \checkmark &  & \checkmark & & \checkmark &
\checkmark& \checkmark & U\\
API Doc. &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
Test Plan &  & \checkmark &  &  & & \checkmark & & & &  U\\
Test Cases & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & U\\
\bottomrule
\end{tabular}
\caption{Comparison of Recommended Artifacts in Software Development Guidelines
to Artifacts in LBM Projects (C for Common, U for Uncommon and R for Rare)}
\label{Tbl_Guidelines}
\end{center}
\end{table}

@@

\item \textbf{API documentation} shows programmers the services or data provided
by the software application (or library) through such resources as its methods
or objects \citep{MengEtAl2018}.  API documentation has been found to be a
critical factor for API understandability (Section~\ref{softwarequalities})
\citep{MengEtAl2018}. Some API documentation can be generated using tools like
Doxygen, but the when learning a new API developers mostly use code examples and
tutorials \citep{MengEtAl2018}.  API documentation was rare within our sample of
LBM packages.

@@

Development tools support the development of end products, but do not become
part of them, unlike dependencies that remain in the application once it is
released \citep[p.\ 506]{GhezziEtAl2003}. Although not shown in
Table~\ref{tbl_tools}, debuggers were also likely used.  Only three (ESPResSo,
Ludwig and Musubi) packages mentioned CI tools, like Travis
CI. Code editors and compilers were explicitly noted to have been used by
several packages, and were likely used by all of them. One of the packages
(Ludwig) explicitly noted the use of proprietary unit testing code written in C.
Likewise, the use of proprietary code for verifying the correctness of output
was noted by one of the developers (pyLBM). Similar tools were likely used when
developing the other software packages.

@@

\subsection{Version Control Usage Compared to Research Software in General}

@@

\subsection{CI Usage Compared to Research Software in General}

As mentioned in Section~\ref{SecSurfCorrectAndVerifiab}, CI is rarely used in
LBM software (3 of 24 packages or 12.5\%), although interviews with developers
suggest that the actual rate might be higher. Low use of CI for LBM contrasts
with how frequently it is recommended in research software development
guidelines
\citep{BrettEtAl2021, Brown2015, ThielEtAl2020, Zadka2018, vanGompelEtAl2016}
and its popularity with non-research software. \citet{HiltonEtAl2016} show CI
usage across a large sample of GitHub projects (34,544 projects) is at least
40\%, with more popular projects (as measured by stars), having even higher
usage rates. Moreover, \citet{HiltonEtAl2016} predict that CI adoption rates for
open-source projects will increase even further in the future.  We could not
find published data on the CI usage rates for research software in general.
However, we saw 17\% utilization in the specific domain of imaging software
\citep{Dong2021}, which is comparable to the LBM usage rate. Our impression is
that medical imaging software and LBM software are not unique -- most research
software domains likely lag behind the best practice recommendations, and
non-research software usage rates, for continuous integration.

CI is the process of building and testing software on every push to the code
repository, with pushes done very frequently \citep[p.\ 13]
{HumbleAndFarley2010}, \citep{ShahinEtAl2017, Fowler2006}.  CI consists of the
following elements:
\begin{itemize}
	\item A version control system \citep{Fowler2006}. To be effective, all
	files should be under version control, not just code files.  Anything that
	is needed to build, install and run the software should be under version
	control, including configuration files, build scripts, test harnesses, and
	operating system configurations \citep[p. 19]{HumbleAndFarley2010}.
	Fortunately for the LBM domain, as discussed above, most alive projects
	(73\% of those assessed) use version control.
	\item A fully automated build system \citep{Fowler2006}.  As \citet[p.\
	5]{HumbleAndFarley2010} point out, deploying software manually is an
	anti-pattern.  Fortunately, as discussed above, most LBM projects use build
	automation.
	\item An automated test system \citep{Fowler2006}. Building quality software
	involves creating automated tests at the unit, component, and acceptance
	test level, and executing these tests whenever a change is made to the code,
	its configuration, the environment, or the software stack that it runs on
	\citep[p.\ 83]{HumbleAndFarley2010}. As Table~\ref{artifactspresent} shows
	test cases are in the uncommon category for LBM software artifacts, which
	means that some LBM projects will need to increase their testing automation
	if they wish to pursue CI.  The usual advice for CI is to keep the build and
	test process short \citep[p.\ 60]{HumbleAndFarley2010}. Given that LBM can
	be computationally expensive, the tests run with every check-in may need to
	focus on simple code interface tests, saving larger more complex tests for
	less frequent runs.  A more sophisticated option for when tests create a
	bottleneck for merges is CIVET (Continuous Integration,
	Verification, Enhancement, and Testing), which solves this problem by
	intelligently pinning, canceling, and if necessary, restarting jobs as
	merges occur, so that two merges are being tested at all times
	\citep{SlaughterEtAl2021}.
	\item An automated system for other tasks, such as code checking,
	documentation building and web-site updating.  These other tasks are not
	essential parts of CI, but they can be incorporated to improve the quality
	of the code and the communication between developers and users. For
	instance, a static analysis (possibly via linters) of the code may find poor
	programming practice or lack of adherence to adopted coding standards.
	\citet{SlaughterEtAl2021} provides another example of an automated task --
	checking that a test specification includes the test's motivation, a test
	description, and a design description for all changes. 
	\item An integration build system to pull everything together.  Every time
	there is a check-in (for instance a pull request), the integration server
	automatically checks out the sources onto the integration machine, starts a
	build, runs tests, and informs the committer of the results. 
\end{itemize}

Although CI can take some time and effort to set-up and integrate into a team's
workflow, the benefits can be significant, as follows:

\begin{itemize}
	\item Headaches associated with a separate integration phase are eliminated
	\citep{Fowler2006}, \citep[p.\ 20]{HumbleAndFarley2010}. If integration of
	the work of different developers, or even separate chunks of work by the
	same developer, is postponed, integration problems are inevitable.  By
	continuously integrating problems are immediately obvious, and the source of
	the problem can be isolated to the small increment that was just committed.
	\item Bugs can be quickly detected and removed \citep{Fowler2006} via
	automated testing.  To improve productivity, defects are best discovered and
	fixed at the point where they are introduced \citep[p.\
	23]{HumbleAndFarley2010}.  Errors are not just found in the code, but also
	in the files and scripts related to configuration management \citep[p.\
	18]{HumbleAndFarley2010}.
	\item Developers are always working on a stable base, since the master
	branch will always be working.  A stable base passes all tests and, if the
	CI system is configured with generators and linters, it will also have up to
	date documentation and coding standard compliant code.  A stable base
	improves developer productivity, allowing them to focus on coding, testing
	and documentation.
\end{itemize}

Although there are still challenges, setting up a CI system has never been
easier than it is today.  The option of installing a dedicated CI server (either
physically or virtually) exists with tools such as
\href{https://www.jenkins.io/} {Jenkins}, \href{http://buildbot.net/}
{Buildbot}, \href{https://www.gocd.org/} {Go}, and
\href{http://integrity.github.io/} {Integrity}. However, installation on your
own server is not necessary since there are many hosted CI tool solutions
available, such as: \href{https://travis-ci.org/} {Travis CI},
\href{https://github.com/features/actions} {GitHub Actions} and
\href{https://circleci.com/} {CircleCI}.  Getting started with a hosted CI is
straightforward.  All that is required to begin is editing a few lines of a
simple YAML configuration file in the project's root directory.

Challenges that exist for adopting CI include lack of awareness and
transparency, coordination and collaboration challenges, lack of expertise and
skills, more pressure and workload for team members, general resistance to
change, scepticism and distrust on continuous practices \citep{ShahinEtAl2017}.
The most common reason given for not adopting CI is that developers on my
project are not familiar enough with CI \citep{HiltonEtAl2016}.  These problems
can be mitigated via planning and documentation, promoting a team mindset,
adopting new rules and policies, improving testing activities, and decomposing
development into smaller units \citep{ShahinEtAl2017}.

@@

Interviews with developers confirmed a similar project management process. In
teams of only a couple of developers, additions of new features or major changes
are discussed with the entire team. Projects with more than a couple developers
have lead developer roles. These lead developers review potential additions to
the software. One of the developers (ESPResSo) noted that an ad hoc peer review
process is used to assess major changes and additions. Using peer review (also
called technical review) matches with recommended practice for research software
\citep{HerouxEtAl2008, Givler2020, OrvizEtAl2017, USGS2019}.

@@

\subsection{Design Principles and Priorities} \label{Sec_DesPrincipAndPriorities}

@@

\subsection{Documentation as Part of the Development Process}

@@

Although Table~\ref{Tbl_Guidelines} shows that design documentation is
frequently suggested, the recommendations are often short on details. For
instance, the guidelines usually encourage modular design, but are generally
silent on what criteria to use for the modularization.  Little information is
given on how to document the design itself. Some non-mutually exclusive ideas
for documenting the design, roughly in order of increasing effort, are as
follows:

\begin{itemize}
	\item Explicitly state the design goals and priorities, so that those using
	and developing your software know part of the rationale for the design
	decisions. Sailfish is an example where this has been done, as discussed in
	Section~\ref{Sec_DesPrincipAndPriorities}.
	\item Write down the likely changes that the design is intended to support.
	Since most designs are implicitly (or explicitly) motivated to facilitate
	future change, recording the likely changes can help readers understand the
	design.
	\item Explicitly record the philosophy or principle behind the
	decomposition. One potential principle, related to the previously mentioned
	list of likely changes, is the principle of information hiding
	\citep{Parnas1972a}.  This principle supports design for change through the
	``secrets'' of each module. A text blurb can be given for each module or
	class to describe its secrets and the services it provides.
	\item Rigorously document the design following the template for a Module
	Guide (MG) \citep{ParnasEtAl1984}.  An MG organizes the modules in a
	hierarchy by their secrets. The application of the Parnas approach to
	scientific software has been illustrated by applying it to the example of a
	mesh generator \citep{SmithAndYu2009}.
	\item Graphically represent the design using Unified Modelling Language
	class diagrams.  This approach is suited to object oriented design and
	designs that use patterns \citep{Gamma1995}.
	\item Explain the design using data flow diagrams to show typical use cases
	for how inputs are transformed.
	\item Using a textual description, list for each module (or class), the
	state variables (if any), exported constants and all exported access
	programs.  This shows the interface that can used to access each module's
	services.
	\item Document the Module Interface Specification (MIS)
	\citep{HoffmanAndStrooper1995}. An MIS is an abstract model that 
	formally shows the state variables, inputs and outputs
	\citep{ElSheikhEtAl2004, SmithAndYu2009}.
\end{itemize}

@@

	\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
	\textbf{Technology Hurdles} Technology pain points for LBM developers
	include setting up parallelization and CI. The pain
	point survey of \citep{WieseEtAl2019} also highlighted technical-related
	problems like dependency management, cross-platform compatibility,
	CI, hardware issues and operating system issues.

@@

Some LBM developers improve reproducibility via automating the verification
tests and making them part of a CI process (mentioned in
Section~\ref{Sec_CompareTools}).  Providing unit tests improves reproducibility
because future developers will require a means to verify that they can obtain
the same computational results as today.  Unit tests are a good way to embed
self-diagnostics of reproducibility in the code \citep{BenureauAndRougier2017}.

@@

In this section we provide recommendations to address the pain points from
Section~\ref{painpoints} to answer~\rqref{RQ_Recommend}.  Our recommendations
are not lists of criticisms for what should have been done in the past, or what
should be done now; they are just suggestions for consideration in the future.
We will not be repeating ideas here that have been discussed in previous
sections, such as CI, documentation of APIs, etc.  Our aim
is to mention ideas that are at least somewhat beyond conventional best
practices. The ideas listed here have the potential to become best practices in
the medium to long-term. The ideas are presented roughly in the order of
increasing novelty.

@@

With the exception of \citet{ThielEtAl2020}, the research software guidelines
that we consulted do not mention linters.  However, we feel linters have the
potential to improve code quality at a relatively low cost.  A linter is a tool
that statically analyzes code to find programming errors, stylistic
inconsistencies and suspicious constructs \citep{Wikipedia2022_Lint}. Linters
can be used to spot check code files, or even better as part of a continuous
integration system, as discussed in Section~\ref{Sec_CompareTools}.

@@

Our threats to validity categories come from an analysis of software engineering
secondary studies by \citet{AmpatzoglouEtAl2019}.  We use the classification of
\citep{AmpatzoglouEtAl2019} because studying a collection of software is a form
of secondary study, where a secondary study analyses the data from a set of
primary studies.  A common example of a secondary study is a systematic
literature review.  Our methodology is essentially a systematic software review
-- the primary studies are the software, and our work collects and analyzes
these primary studies.

\subsection{Reliability}

A study is reliable if its data and analysis are independent of the specific
researcher(s) doing the study.  That is, repetition of the study by a different
researcher should achieve the same results, as long as the same methodology is
used \citep{AmpatzoglouEtAl2019}.  For our study the identified reliability
related threats include the following:

\begin{itemize}
\item In our methodology the manual measures are done by one individual for all
packages. The measures may be influenced by the graders abilities, experience
and biases.  
\item Completing the manual measures of all packages took place over several
months. During this time the contents of the software repositories may change
and the reviewer's judgment may drift.
\end{itemize}

We investigated the risk of reviewer dependence in previous work
\citep{SmithEtAl2016}, where we demonstrated that the measurement process is
reasonably reproducible by having five products graded by a second reviewer. The
ranking via this independent review was almost identical to the original
ranking. As long as each grader uses consistent definitions, the relative
comparisons in the AHP results will be consistent between graders.

\subsection{Construct Validity}

Construct validity is threatened if metrics are used that do not measure what
they are intended to measure \citep{AmpatzoglouEtAl2019}. The construct
validity related threats for the LBM study include the following:

\begin{itemize}
\item Our approximation of popularity by stars and watches
(Section~\ref{repmetrics}) may not be valid. 
\item Our overall AHP ranking makes the unrealistic assumption of equal
weighting between qualities (Section~\ref{Sec_OverallQuality}).
\item Our measure of surface robustness in only based on two pieces of data per
project. The measurement of robustness could be expanded, as it currently only
measures the impact of unexpected input. Other intentional faults could be
introduced, but this would require a larger investment of measurement time. 
\item Reusability (Section~\ref{reusabilityresults}) is assessed via the number
of code files and LOC per file. While this measure is indicative of modularity,
it is possible that some packages have many files, with few LOC, but the files
do not contain source code that is easily reusable. The files may be poorly
formatted, or the source code may be vague and have ambiguous identifiers. 
\item Maintainability may be incorrectly estimated, since it is assumed to be
improved by a higher ratio of comments to source code
(Section~\ref{Sec_Maintainability}). 
\item The measurement of understandability
(Section~\ref{Sec_SurfUnderstandability}) relies on 10 random source code files.
By random chance, the 10 files that were chosen may not be representative. 
\item We measure our software qualities indirectly, rather than directly.
Unfortunately meaningful direct measures for qualities like maintainability,
reusability and verifiability, are not available.  Therefore, we follow the
usual approach of assuming that following procedures and adhering to standards
yields a higher-quality product \citep[p.\ 112]{VanVliet2000}.
\end{itemize}

\subsection{Internal Validity} \label{Sec_InternalValidity}

A study is internally valid if the discovered causal relations are trustworthy
and cannot be explained by other factors \citep{AmpatzoglouEtAl2019}. The
internal validity threats include the following:

\begin{itemize}
\item We may have missed a relevant software package in our search for LBM
software packages (Section~\ref{identifysoftware}).
\item A temporal gap exists between the manual measurement of Musubi and the
measurement of all other packages (Section~\ref{identifysoftware}). This gap
caused a mismatch between the time of manual and automated data collection
(Section~\ref{repmetrics}), since we updated the automated measures (like the
number of files, LOC, etc.) when Musubi was added.
\item Our interviews were with a relatively small sample of four LBM developers.
The pain points they expressed may not be representative of others in the
community.
\item In our manual measures of the software repositories some data may not have
been measured due to technology issues like broken links. This issue arose with
the measurement of Palabos, which had a broken link to its user manual, as noted
in Section~\ref{Sec_CompareArtifacts}. 
\item Our methodology assumes that artifacts will be present in the repositories
to approach used for software development.  However, not all activities will
necessarily leave a trace in the repositories. For instance, the use of unit
testing and CI was mentioned in the artifacts of only three (ESPResSo, Ludwig,
Musubi) packages (Section~\ref{SecSurfCorrectAndVerifiab}). However, interviews
suggested a more frequent use of both unit testing and CI. For example, OpenLB,
pyLBM, and TCLB use such methods during development despite this not being
explicit from an analysis of the material available online.
\end{itemize}

Although we found that unit testing and CI were used more often than the
repository artifacts suggested, we did not change the manual measurement
results.  Given that we could not interview a representative for every project,
we left the data as collected to keep our measures consistent.  Moreover, CI and
unit testing done in private should not be counted as equivalent to the public
versions we detected.  When CI and unit testing are done outside of the public
repository, they are not as effective, since they are not available for others
to inspect and reproduce.

\subsection{External Validity}

External validity concerns whether the results of the study can be generalized
(applied) to other situations/cases \citep{AmpatzoglouEtAl2019}.  The scope of
our work is LBM solvers.  We are confident that our search was exhaustive and
that we did not miss any highly popular examples.  Therefore, our validity
concerns about LBM software observations are internal
(Section~\ref{Sec_InternalValidity}), not external.  However, our hope is that
the trends observed and the lessons learned for LBM software can be applied to
other research software.  With that in mind we identified the following threat
to external validity:

\begin{itemize}
\item The development of LBM software is fundamentally different from other
research software.
\end{itemize}

@@

TO do
- check references
- maybe add a continuous integration diagram?
